---
title: "CS598 - Project 1"
author: "Xiaoming Ji"
output: pdf_document
---
# Computer System

## Hardware
- Dell Precision Tower 5810
- CPU: Intel Xeon E5-1607 @ 3.10GHz
- Memory: 32GB

## Software
- OS: Windows 10 Professional 64bit
- R: 3.5.1
- R Packages:
    - forecast_8.4
    - tidyverse_1.2.1
    - lubridate_1.7.4

## Models

3 approaches (total 4 models) are used to generate the prediction:

- Naive model
- Seasonal naive model
- Dynamic model: for fold 1 to 6, regression model (tslm) is used. Start from fold 7, since the training data has more than 2 years of data, STL+ARIMA (method='arima', ic='bic') model is built to make the prediction.

## Missing value handling

- Weekly_Sales: replace missing value with 0. 
- IsHoliday: search through the training data to find the IsHoliday of same date. See function: `fill_missing_holiday`

*Note*: my testing show more sophisticated imputation approach won't improve the performance.

## Test results
```{r, message=FALSE, warning=FALSE, include=FALSE}
mypackages = c("forecast", "tidyverse", "kableExtra", "knitr")   # required packages
tmp = setdiff(mypackages, rownames(installed.packages()))  # packages need to be installed
if (length(tmp) > 0) install.packages(tmp)
lapply(mypackages, require, character.only = TRUE)
```

```{r, echo=FALSE, message=FALSE}
source("mymain.R")

# read in original train / test dataframes
train_all <- readr::read_csv('train_origin.csv')

# read in train / test dataframes
train <- readr::read_csv('train.csv')
test <- read_csv('test.csv', col_types = list(
    Weekly_Pred1 = col_double(),
    Weekly_Pred2 = col_double(),
    Weekly_Pred3 = col_double()
))

# save weighted mean absolute error WMAE
num_folds <- 10
wae <- tibble(
    model_one = rep(0, num_folds), 
    model_two = rep(0, num_folds), 
    model_three = rep(0, num_folds)
)
```

```{r message=TRUE, message=FALSE, warning=FALSE, include=FALSE}
# DEV.STORES = c(1)
# DEV.DEPARTMENTS = c(1, 3, 5, 10, 13, 21, 22)
# 
# train = train[train$Store %in% DEV.STORES & train$Dept %in% DEV.DEPARTMENTS, ]
# test = test[test$Store %in% DEV.STORES & test$Dept %in% DEV.DEPARTMENTS, ]

start_time = proc.time()
# time-series CV
for (t in 1:num_folds) {
  cat("Fold:", t, "\n")
  # *** THIS IS YOUR PREDICTION FUNCTION ***
  mypredict()
  
  # Load fold file 
  # You should add this to your training data in the next call 
  # to mypredict()
  fold_file <- paste0('fold_', t, '.csv')
  new_test <- readr::read_csv(fold_file)
  # new_test = new_test[new_test$Store %in% DEV.STORES & new_test$Dept %in% DEV.DEPARTMENTS, ]

  # extract predictions matching up to the current fold
  scoring_tbl <- new_test %>% 
      left_join(test, by = c('Date', 'Store', 'Dept'))
  
  # compute WMAE
  actuals <- scoring_tbl$Weekly_Sales
  preds <- select(scoring_tbl, contains('Weekly_Pred'))
  weights <- if_else(scoring_tbl$IsHoliday.x, 5, 1)
  wae[t, ] <- colSums(weights * abs(actuals - preds)) / sum(weights)
  
  print(wae[t,])
}

computation.time = print((proc.time() - start_time)[3])

# save results to a file for grading
#readr::write_csv(wae, 'Error.csv')
colMeans(wae)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
test.results = cbind(Fold= c(rep(1:num_folds), 'Average'), rbind(wae, colMeans(wae)))
kable(test.results, col.names = c('Fold', 'Naive','SNaive','Dynamic')) %>%
  kable_styling(bootstrap_options = c("striped"), full_width = F) 
```

Computation time: `r computation.time` seconds
